### 2018-09-23 paper/complexity.tex
New upper bound and new proof for maximal number of hitting sets, based on Sperner's Theorem.

% -------------------------------------------------------------------------
\subsection{Estimation of $\ell$}
The first factor $\ell$
has worst-case upper bound $\ell\le |\Sigma|^k$. As an example, where this upper bound 
is really met, consider the reference process
\[
RUN(\Sigma) = e:\Sigma \then RUN(\Sigma).
\]
The normalised transition graph of this process has a single state, and its initials
are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(k)$ can 
never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly 
$|\Sigma|^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
for each of their events.

% -------------------------------------------------------------------------
\subsection{Estimation of $h$}
Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
where all $A_i$ are disjoint. In this case,  $h~~=~~|\minhits(P/s)|~~=~~
\prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
that we prove in the following lemma.
%
\begin{lemma}\label{lemma:minhitmax}
Let $n\ge 2$ be any positive integer. We call a sequence
%
$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$
%
a partition of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$. Define
%
$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
\alpha\in \mathbb{N}\big\}.$$
%
\end{lemma}
\begin{proof}
Let $n\ge 2$. Obviously, \begin{equation}\label{eq:1} n\le
p(n).\end{equation} Define $\pi(n)$ to be the following partition  of $n$:
%
$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
%
Then the product of the numbers in $\pi(n)$ equals  $p(n)$. In the remainder
of the proof we have to show that $p(n)$ is really maximal in the sense of
the lemma.

Since the number of partitions of a given $n$ is finite,
%
$$\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$$
%
exists. Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with
maximal product. Then, obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then
from~(\ref{eq:1}) we have
$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a
product value which is not less then the original partition $a_1, \ldots ,
a_{\alpha}$. This process can be repeated, until every $a_i$ in the resulting
partition is either 2 or 3. Hence there exists a partition of $n$ with
maximal product, such that every term is $2$ or $3$. Since $6=2+2+2=3 + 3$
and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than
$3$. Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\},
i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$
be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$
with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have

\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
 k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
This proves the lemma.
\xbox
\end{proof}
%
From Lemma~\ref{lemma:minhitmax}, we conclude that
$$h=\begin{cases}
3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\  |\Sigma|~~\equiv~~0 \mod 3\\
3^{\lfloor\frac{|\Sigma|}{3}\rfloor}  + 3^{\lfloor\frac{|\Sigma|}{3}\rfloor-1}&\text{if}\ |\Sigma|~~\equiv~~1\mod 3\\
2\cdot 3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\ |\Sigma|~~\equiv~~2 \mod 3
\end{cases}
$$
%
We note that $h$ is significantly smaller than $2^{|\Sigma|}-1$, this can
also be intuitively motivated by the fact that  for every $H\in
\minhits(P/s)$, all of its true subsets and true supersets are {\it not}
contained in $\minhits(P/s)$.

% -------------------------------------------------------------------------
\subsection{Upper Bounds of Test Executions for Checking Failures Refinement}
According to  Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
$k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
we use the formula for the sum of the geometric progression.
%
\[
h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
 \text{or, asymptotically,}\  O(3^{\lfloor\frac{|\Sigma|}{3}\rfloor}\cdot|\Sigma|^{(pq-1)}).
 \]
%
From Lemma~\ref{lemma:hseta} we know that the worst-case bound for $h$
above cannot be further improved, since the full collection of minimal
hitting sets needs to be checked to verify $conf$. The
authors of~\cite{Hennessy:1988:ATP:50497} and
\cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
subset of $\Sigma$ whose events cannot be completely refused in a given
process state of the reference model; this leads to a worst-case estimate of
$2^{|\Sigma|}-1$ for the number of different sets to be offered to the SUT in
the last step of the test execution, so our approach reduces the number of
test executions in comparison
to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.

@@

According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
refinement just contains the adaptive test case $U_T(pq-1)$. As derived for $U_F(k)$
above, the number of executions performed by $(Q\parallel[\Sigma] U_R(pq-1))$ 
is bounded by $|\Sigma|^{pq-1}$.

>>>>

% -------------------------------------------------------------------------
\subsection{Estimation of $\ell$}
The first factor $\ell$
has worst-case upper bound $\ell\le \card{\Sigma}^k$. As an example, where this upper bound 
is really met, consider the reference process
\[
RUN(\Sigma) = e:\Sigma \then RUN(\Sigma).
\]
The normalised transition graph of this process has a single state, and its initials
are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(k)$ can 
never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly 
$\card{\Sigma}^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
for each of their events.

% -------------------------------------------------------------------------
\subsection{Estimation of $h$}
Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
acceptances, the cardinality   $h = \minhits(P/s)$ has the upper bound
\[
h \le \binom{n}{\lfloor\frac{n}{2}\rfloor}, \quad \text{where $n = \card{\Sigma}$.}
\]
This follows from Theorem~\ref{th:sperner} and the fact that $\minhits(P/s)$ is a 
Sperner Family, as introduced in Section~\ref{sec:hit}. The next theorem shows that this
upper bound can really be reached by collections of minimal hitting sets.

\begin{theorem}
\label{th:upperboundh}
Let $n\ge 2$ be any positive integer and $\Sigma$  any finite set of cardinality $n$.
Let $C\subseteq 2^\Sigma$ be any collection of  subsets of $\Sigma$. Then  
\[
\text{max}\{\card{\text{minHit}(C)}~|~C\subseteq 2^\Sigma\}=\binom{n}{\lfloor\frac{n}{2}\rfloor}
\]
\end{theorem}
\begin{proof}
For any subset  $B$  of $\Sigma$, define  $A_{B}:=\Sigma\setminus B$.
Define 
$$
C=\{A_{B}~|~ B\subseteq \Sigma, \card{B} < \lfloor\frac{n}{2}\rfloor\}.
$$
Let $H$ be any minimal Hitting set of $C$. Then $H$ contains at least $\lfloor\frac{n}{2}\rfloor$ elements, because 
otherwise there existed an $A_{H}\in C$ with $A_H\cap H=\varnothing$. 
Since $\card{A_{B}}=n-\card{B}>n-\lfloor\frac{n}{2}\rfloor$, for any $B\subseteq \Sigma$ with $\card{B}< \lfloor\frac{n}{2}\rfloor$, we conclude that
 any  subset of $\Sigma$  with at least $\lfloor\frac{n}{2}\rfloor$ elements 
 has a non-empty intersection with all  $A_B\in C$.  
 The number of subsets in $\Sigma$ with exactly $\lfloor \frac{n}{2}\rfloor$ elements
 is $\binom{n}{\lfloor\frac{n}{2}\rfloor}$.
Thus $\card{\text{minHit}(C)} \ge \binom{n}{\lfloor\frac{n}{2}\rfloor}$, and 
Theorem~\ref{th:sperner} yields $\card{\text{minHit}(C)} = \binom{n}{\lfloor\frac{n}{2}\rfloor}$.
\xbox
\end{proof}

To get an approximation of the maximal size of $\minhits(P/s)$ for large $n$, recall
Stirling's approximation~\cite[p.~112]{Graham:1994:CMF:562056}
\[
m! \approx \sqrt{2\pi m} \cdot \big( \frac{m}{e} \big)^m, \quad\text{where $e$ denotes the Euler Number.}
\]
Applying this approximation to the   case where $n = \card{\Sigma}$ is even and $m = (n/2)$  
results in
\[
h \le \binom{2m}{m} \approx \frac{4^m}{\sqrt{\pi m}} = 
\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}.
\]
%
We note that, due to this approximation result, 
$h$ is   smaller than $2^{\card{\Sigma}}-1$, the cardinality of the 
non-empty subsets of $\Sigma$.

% -------------------------------------------------------------------------
\subsection{Upper Bounds of Test Executions for Checking Failures Refinement}
According to  Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
$k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
we use the formula for the sum of the geometric progression.
%
\[
h\cdot \big( \frac{1-\card{\Sigma}^{pq}}{1-\card{\Sigma}} \big),\
 \text{or, asymptotically,}\  O(\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}\cdot\card{\Sigma}^{(pq-1)}).
 \]
%
From Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
above cannot be further improved, since the full collection of minimal
hitting sets needs to be checked to verify $conf$. The
authors of~\cite{Hennessy:1988:ATP:50497} and
\cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
subset of $\Sigma$ whose events cannot be completely refused in a given
process state of the reference model; this leads to a worst-case estimate of
$2^{\card{\Sigma}}-1$ for the number of different sets to be offered to the SUT in
the last step of the test execution, so our approach reduces the number of
test executions in comparison
to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} by a factor of
$1/{\sqrt{2\pi \card{\Sigma}}}$.

In Fig.~\ref{fig:minhita}, the reduction is visualised by a function plot of
$2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.

% .....................................................................................
 \begin{figure}
 %%\hspace*{-40mm}
 \begin{center}
\includegraphics[width=.8\textwidth]{minhit-fig.pdf}
\end{center}
\vspace*{-10mm}
\caption{Function plot $2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.}
 \label{fig:minhita}
 \end{figure}
% .......................................................................................

@@

According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
refinement just contains the adaptive test case $U_T(pq-1)$. As derived for $U_F(k)$
above, the number of executions performed by $(Q\parallel[\Sigma] U_R(pq-1))$ 
is bounded by $\card{\Sigma}^{pq-1}$.

### 2018-09-23 paper/preliminaries.tex
New upper bound and new proof for maximal number of hitting sets, based on Sperner's Theorem.

>>>>

In preparation of complexity results to be presented in Section~\ref{sec:complexity}, we consider the question how many minimal hitting sets can maximally exist for a given 
collection of finite sets. To this end, the following definitions and results are 
useful\footnote{The authors are grateful to @todo for suggesting the applicability of Sperner's Theorem in the context of the work presented here.}. 
A \emph{Sperner Family} is a collection $S \subseteq 2^\Sigma$ of sets from a given universe $\Sigma$ that do not contain each other, that is, 
$H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq H_2\in S$. 
Specialising   antichains known from   partial orders to finite sets partially ordered by $\subseteq$ results in Sperner families.

We observe that the collection of minimal hitting sets constructed from the minimal acceptances of a CSP process state build a Sperner Family. Therefore, the following theorem can be applied.
\begin{theorem}[Sperner's Theorem~\cite{sperner_satz_1928}]
\label{th:sperner}
The largest Sperner family $S$ 
over an $n$-element universe $\Sigma$ has cardinality 
$$
\card{S} = \binom{n}{\lfloor\frac{n}{2}\rfloor}.
$$
\xbox
\end{theorem}
It will be shown in Section~\ref{sec:complexity} that this upper bound can actually be reached by hitting sets associated with the minimal acceptances of a CSP process state.
